---
title: "Wikipedia History Proof of Concept"
author: "Adam Campbell"
date: "25/11/2020"
output: html_document
---

The goal of this script is to download data from wikipedia on the complete edit history of an article.  Eventually I might like to turn this into a R package, but for the time being I'll keep this as a markdown file for a proof of concept.

To do this work I made use [Wikipedia's API Tool](https://www.mediawiki.org/wiki/API:Revisions#Example_2:_Get_last_five_revisions_of_a_page_filtered_by_date_and_user) and [Dataquest's API tool](https://www.dataquest.io/blog/r-api-tutorial/)

First load the libraries for doing the query.
```{r message=FALSE, load-libraries}
library(httr)
library(jsonlite)
library(tidyverse)
library(lubridate)
```

Here are two functions, `get_500_recent` will get the 500 (Wikipedia's limit) of most recent edits on the page,  `get_500_recent` get's the full edit history.   
```{r}
# get 500 results

get_500_recent <- function(title, start_date = "start"){
  base_list = list(
    action = "query", 
    prop = "revisions",
    rvprop = "timestamp",
    rvlimit = "500",
    rvslots = "main",
    formatversion = "2",
    rvdir = "newer",
    format = "json"
  )
  
  # make list
  if (start_date == "start") {
    supp_list = list(
      titles = title
    )  
  } else {
    supp_list = list(
      titles = title,
      rvstart = start_date
    )
  }

  query_list = c(base_list, supp_list)
  
  res = GET(
    "http://en.wikipedia.org/w/api.php", 
    query = query_list
    )

  data = fromJSON(rawToChar(res$content))
  
  dom <- data$query$pages$revisions[[1]]
  
  return(dom)
}

# get full edit history

get_edit_history <- function(title){
  
  first_data <- get_500_recent(title)
  
  if (nrow(first_data) < 500 ) {
    return(first_data)
  } 
  
  fetched_data <- first_data
  data_out <- fetched_data
  
  while (nrow(fetched_data) == 500 ){
    data_out <- data_out %>% filter(row_number() <= n()-1)
    date_start <- as.character(tail(fetched_data, n=1))
    fetched_data <- get_500_recent(title, date_start)
    data_out <- rbind(data_out, fetched_data)
  }
  return(data_out)
}
```

Let's do a test for "Coronavirus disease 2019"
```{r}
title = "Coronavirus_disease_2019"
my_data <- get_edit_history(title)
```

Have to formulate the dates
```{r}
my_data$dates <- as_date(my_data$timestamp)
my_data$testdatetime <- parse_date_time(my_data$timestamp,"y m d HMS")
my_data$testmonth <- month(my_data$timestamp)
```

And do a simple plot of the results
```{r}
ggplot(my_data) + aes(x = lubridate::floor_date(dates, "month")) + 
  geom_bar() +
  theme_minimal()
```

